---
title: "SOW 1 - Data Acquisition"
format:
  html:
    embed-resources: true
    toc: true
    code-fold: true
execute:
  message: false
  warning: false
editor: visual
editor_options: 
  chunk_output_type: console
---

# Overview

# Data Sources

1.  Census

    -   PUMS
    -   ACS
    -   Census Bureau LODES data
        -   is this at county level
        -   is at state level

2.  [EEOC](https://www.eeoc.gov/)

    -   Abbas hasn't found this available in R yet
    -   

3.  Bureau of Labor Statistics (BLS)

    -   blsAPI (Google these)

    -   blsrapeR

## Load packages

```{r}
#| message: false
library(tidycensus)
library(tidyverse)
library(shiny)
```

Here we will extract the variables of interest. Those variables include:

1.  Age

2.  Income

3.  Race

4.  Sex

```{r}
vars_2021 <- load_variables(year="2021", dataset="acs5")

vars_2021 |> 
  filter(str_detect(label, "Household income")) |> 
  slice(1:5) |>
  pull(label)
  
```

For the our first pass, let's see if we can look at NY, specifically Monroe County.

```{r}
ny_pop <- get_acs(geography = "county", 
                     variables = "B01003_001", 
                     state = "NY",
                     geometry = TRUE) 

ny_pop 
```

Population

```{r}
library(leaflet)
library(stringr)
library(sf)

pal <- colorQuantile(palette = "viridis", domain = ny_pop$estimate, n = 10)

ny_pop %>%
    st_transform(crs = "+init=epsg:4326") %>%
    leaflet(width = "100%") %>%
    addProviderTiles(provider = "CartoDB.Positron") %>%
    addPolygons(popup = ~ str_extract(NAME, "^([^,]*)"),
                stroke = FALSE,
                smoothFactor = 0,
                fillOpacity = 0.7,
                color = ~ pal(estimate)) %>%
    addLegend("bottomright", 
              pal = pal, 
              values = ~ estimate,
              title = "Population percentiles",
              opacity = 1)
```

Median home value

```{r}


median_home_map <- function(county){
  roc_ny_value <- get_acs(geography = "tract", 
                    variables = "B25077_001", 
                    state = "NY",
                    county = county,
                    geometry = TRUE)

pal <- colorNumeric(palette = "viridis", 
                    domain = roc_ny_value$estimate)

roc_ny_value %>%
    st_transform(crs = "+init=epsg:4326") %>%
    leaflet(width = "100%") %>%
    addProviderTiles(provider = "CartoDB.Positron") %>%
    addPolygons(popup = ~ str_extract(NAME, "^([^,]*)"),
                stroke = FALSE,
                smoothFactor = 0,
                fillOpacity = 0.4,
                color = ~ pal(estimate)) %>%
    addLegend("bottomright", 
              pal = pal, 
              values = ~ estimate,
              title = "Median Home Value",
              labFormat = labelFormat(prefix = "$"),
              opacity = 1)
}


median_home_map("Erie County")

```

```{r}
# need to look at PUMS vars

pums_vars_2018 <- pums_variables |>
  filter(year == 2018, survey == "acs1")

ny_pums_recoded <- get_pums(
  variables = c("PUMA", "SEX", "AGEP", "SCHL", "CIT"),
  state = "NY",
  survey = "acs1",
  year = 2018,
  recode = TRUE
  )
```

Next steps:

-   Write functions that enable shiny inputs

-   PUMS
